Generalized Tractability

### 2023-04-21 GeneralizedTractability_230418.tex
Updates from Overleaf

\subsection{Notation and fundamental definitions}
Let $T$ be a tractability function defined as follows,
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty),
\end{equation}
A tractability function is a function of simple form that provides an upper bound on the computational complexity of a problem.  Some examples of tractability functions commonly considered are provided in Section \ref{sec:examples}.

@@

@@

Certain common notions of tractability are covered by the general situation described above.
\begin{equation*}
	\begin{array}{r@{\qquad}c@{\qquad}cc}
		& T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		\text{strong polynomial algebraic  tractability} & \max(1,\varepsilon^{-p}) & 1\\
		\text{polynomial algebraic tractability} & \max(1,\varepsilon^{-p})d^{q} & d^{q}\\
		\text{strong polynomial exponential tractability} &  [\max(1,\log(1 + \varepsilon^{-1})]^p & 1\\
		\text{polynomial exponential tractability} &
		[\max(1,\log(1 + \varepsilon^{-1}))]^p  d^{q} & d^{q}
	\end{array}
\end{equation*}
Here above, $\vp = (p,q)$ in some cases. A general reference 
for results on algebraic tractability is \cite{NW08}, for exponential tractability we refer to \cite{HKW19} and \cite{KW19}.

@@

\begin{theorem}\label{thm_main_tract2}
Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions} and \eqref{eq:ptauassume}.  A problem is $T$-tractable iff there exist $\vp\ge \vzero$ and a constant $L_{\vp} > 0$ such that
\begin{equation} \label{eq:tractiff4}
     S_{\vp}:=\sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,d,\vp) \rceil}^\infty \frac{1}{T(\lambda^{-1}_{i,d},d,\vp)}< \infty.
\end{equation}
If \eqref{eq:tractiff4} holds for some $\vp$, let  
\[
\widetilde{\cp}_{\textup{opt}} : = \{\vp^* : \eqref{eq:tractiff4} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty}) \text{ and }  \eqref{eq:tractiff4} \text{ does not hold } \forall \vp \in [\vzero,\vp^*/2)\}.
\]
Then $\cp_{\textup{opt}} \subseteq \widetilde{\cp}_{\textup{opt}}$.
\
\end{theorem}

@@

\begin{example}[Strong polynomial tractability]

Let $ T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty)$
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= (\max\{\varepsilon^{-1},1\})^p .
\]
Then Theorem \ref{thm_main_strong_tract2} yields the iff-condition
\[
 \sup_{d\in\naturals} \sum_{i=L(p)}^\infty (\min\{\lambda_{i,d},1\})^p < \infty.
\]
This essentially recovers the result on strong polynomial tractability in \cite[Theorem 5.1]{NW08}
\end{example}

\begin{example}[Polynomial tractability]
Let $ T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty)$
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= (\max\{\varepsilon^{-1},1\})^p\, d^q.
\]
Then Theorem \ref{thm_main_tract2} yields the iff-condition
\[
 \sup_{d\in\naturals} \sum_{i=L_{\vp} d^{q_1}}^\infty (\min\{\lambda_{i,d},1\})^p d^{-q} < \infty.
\]
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NW08}
\end{example}

\peter{(PETER: Add further examples (e.g., EXP- polynomial tractability, etc.)}
\kachi{\begin{example}[EXP-strong polynomial tractability]
Let $ T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty)$
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= (1+\log\max\{\varepsilon^{-1},1\})^p.
\]
Then Theorem \ref{thm_main_tract2} yields the iff-condition
\[
 \sup_{d\in\naturals} \sum_{i=L_{\vp} d^{q_1}}^\infty (\min\{\lambda_{i,d},1\})^p  < \infty.
\]
\end{example}
\begin{example}[EXP-polynomial tractability]

>>>>

\subsection{Notation and fundamental definitions} \label{sec:not}
Let $T$ be a tractability function defined as follows,
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty),
\end{equation}
A tractability function is a function of simple form that provides an upper bound on the computational complexity of a problem.  Some examples of tractability functions commonly considered are provided in Section \ref{sec:examples}.

@@

Certain common notions of tractability are covered by the general situation described above.
{\small 
\begin{equation*}
	\begin{array}{r@{\qquad}c@{\qquad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		\text{strong polynomial algebraic} & \max(1,\varepsilon^{-p}) & 1\\
		\text{polynomial algebraic} & \max(1,\varepsilon^{-p})d^{q} & d^{q}\\
		\text{strong polynomial exponential} &  [\max(1,\log(1 + \varepsilon^{-1})]^p & 1\\
		\text{polynomial exponential } &
		[\max(1,\log(1 + \varepsilon^{-1}))]^p  d^{q} & d^{q} \\
        \text{quasi-polynomial} & 
        \exp\{p(1+\log(\max(1,\varepsilon^{-p}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}
	\end{array}
\end{equation*}}
Here above, $\vp = (p,q)$ in some cases and $\vp$ is a scalar in other cases. A general reference 
for results on algebraic tractability is \cite{NW08}, for exponential tractability we refer to \cite{HKW19} and \cite{KW19}.

@@

@@

\begin{theorem}\label{thm_main_tract2}
Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions} and \eqref{eq:ptauassume}.  A problem is $T$-tractable iff there exists $\vp\ge \vzero$ and a positive integer $L_{\vp}$ such that
\begin{equation} \label{eq:tractiff4}
     S_{\vp}:=\sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,d,\vp) \rceil}^\infty \frac{1}{T(\lambda^{-1}_{i,d},d,\vp)}< \infty.
\end{equation}
If \eqref{eq:tractiff4} holds for some $\vp$, let  
\[
\widetilde{\cp}_{\textup{opt}} : = \{\vp^* : \eqref{eq:tractiff4} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty}) \text{ and }  \eqref{eq:tractiff4} \text{ does not hold } \forall \vp \in [\vzero,\vp^*/2)\}.
\]
Then $\cp_{\textup{opt}} \subseteq \widetilde{\cp}_{\textup{opt}}$.
\
\end{theorem}

@@

In this section, we illustrate Theorems \ref{thm_main_strong_tract2} and \ref{thm_main_tract2} for the types of tractability given in the table in Section \ref{sec:not}.

\begin{example}[Strong algebraic polynomial tractability]

Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,p)= (\max\{\varepsilon^{-1},1\})^p 
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p \ge 0.
\]
Then Theorem \ref{thm_main_strong_tract2} yields the  equivalent condition for strong algebraic polynomial tractability: 
\[
 \sup_{d\in\naturals} \sum_{i=L_p}^\infty (\min\{\lambda_{i,d},1\})^p < \infty \text{ for some } p \ge 0, L_p \in \naturals.
\]
Furthermore, the optimal $p$ is the infimum of all $p$ for which the above condition holds.
This essentially recovers the result in \cite[Theorem 5.1]{NW08}.
\end{example}

\begin{example}[Algebraic polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,p,q)= (\max\{\varepsilon^{-1},1\})^p\, d^q
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p,q \ge 0.
\]
Then Theorem \ref{thm_main_tract2} yields the  equivalent condition for algebraic polynomial tractability: 
\[
 \sup_{d\in\naturals} \sum_{i=L_{p,q} d^{q}}^\infty (\min\{\lambda_{i,d},1\})^p d^{-q} < \infty \text{ for some } p,q \ge 0, L_{p,q} \in \naturals.
\]
Furthermore, the set of optimal $(p,q)$ is a subset of 
\begin{multline*}
\{(p^*,q^*) : \text{the above condition holds for all }(p,q) \in (p^*,\infty) \times (q^*,\infty) \\
\text{but not for any }(p,q) \in [0,p^*/2) \times [0,q^*/2)\}.
\end{multline*}
This essentially recovers the result on polynomial tractability in \cite[Theorem 5.1]{NW08} \fred{same theorem as above??, what about optimality in \cite[Theorem 5.1]{NW08}}.
\end{example}

\kachi{\begin{example}[EXP-strong polynomial tractability]
Let $ T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty)$
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= (1+\log\max\{\varepsilon^{-1},1\})^p.
\]
Then Theorem \ref{thm_main_tract2} yields the iff-condition
\[
 \sup_{d\in\naturals} \sum_{i=L_{\vp} d^{q_1}}^\infty (\min\{\lambda_{i,d},1\})^p  < \infty.
\]
\end{example}
\begin{example}[EXP-polynomial tractability]
